Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-03-31 GWLDCII.tex
agree with Sian's new version of lem 4.8 (with some small mods)

These $C(W)$ parametrize a subspace of $\Gr(k, n)$ as shown in \cite{wilsonloops}, call it $\Sigma(W)$. The Wilson loop diagrams also define a volume form on $\Sigma(W)$: \bas \Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V_{p_r}} \textrm{d}c_{p_r}}{R(W)} \;. \eas The denominator $R(W)$ is a polynomial defined by $2 \times 2$ and $1 \times 1$ minors of $C(W)$ as defined below.

@@

\begin{lem}\label{lem n-2 bad}
Let $V$ and $W$ be as in Lemma~\ref{lem I}, and suppose that $n-2,n-1 \in I_{n-2}^{V}$.

Then $I_{n-2}^{W}$ and $I_{n-3}^{W}$ contribute the following plusses to the Le diagram of $W$
  \begin{itemize}
  \item An $(n-3)\rightarrow (n-2)$ plus and an $(n-5)\rightarrow (n-1)$ plus.
  \item All of the $I_{n-1}^{V}$ plusses.
  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $n-5\rightarrow n-2$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields an $n-5\rightarrow n-1$ plus instead.
  \item All other plusses of $I_{n-2}^{V}$.
  \item In the Le diagram of $V$, $I_{n-3}^{V}$ gives a plus in the $n-3$ column.  This $+$ is shifted over into the $n-2$ column in $W$.
  \item All other plusses of $I_{n-3}^{V}$.
  \end{itemize}
Furthermore, no element of the Gramann necklace of $V$ gives an $n-5\rightarrow n-1$ plus.
\end{lem}

\begin{proof}
  By Lemma~\ref{lem I} $I_{n-2}^{W}= (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}$.  Also, by the location of $q$ in the WLD,  $n-2\not\in I_{n-3}^{V}$ and and $n-5$ is the index of the lowest vertical edge in $\mathcal{B}$.  Thus this section of the Gramann necklace of $V$ looks like
  \begin{equation}\label{eq necklace}
  I_{n-3}^{V} \underset{\substack{n-3\text{ out}\\n-2\text{ in}}}{\rightarrow} I_{n-2}^{V} \underset{\substack{n-2\text{ out}\\n-5\text{ in}}}{\rightarrow} I_{n-1}^{V} \underset{\substack{n-1\text{ out}\\\text{something in}}}{\rightarrow} I_{n}^{V} \underset{\substack{n \text{ out}\\\text{something in}}}{\rightarrow} I_1^{V}
  \end{equation}
  where the first ``something'' is either $n$ or an element of $I_1^{V}$ and the second ``something'' is an element of $I_1^{V}$.  Additionally all elements not explicitly mentioned must be in $I_1^{V}$ as they remain unchanged through this portion of the necklace.

  Using this information now determine the symmetric difference of $I_{n-2}^{V}$ and $I_1^{V}$: $n-1, n-2$ and possibly $n$ are in $I_{n-2}^{V}$ but not in $I_1^{V}$.  $n-5$ is in $I_1^{V}$ as are at least one and at most two other elements.  If there is one such element call it $a$.  If there are two call them $a$ and $b$ with $a>b$.  This means that the plusses in the Le diagram of $V$ coming from $I_{n-2}^{V}$ are as in the first part of Figure~\ref{fig messy}.  Stepping to $I_{n-1}^{V}$ simply removes the $n-5\rightarrow n-2$ path, see the second part of Figure~\ref{fig messy}.

  \begin{figure}
    \includegraphics{messy}
    \caption{Plusses coming from $I_{n-2}^{V}$ (top left),  $I_{n-1}^{V}$ (top right) and $I_{n}^{V}$ bottom when $n-1, n-2\in I_{n-1}^{V}$.  The blue lines are the non-intersecting paths.  The dashed blue lines may or may not appear, but if one appears then they both do.}\label{fig messy}
  \end{figure}

  Stepping to $I_{n}^{V}$, $n-1$ is taken out and either $n$ is put in if it was not there before, or one of $a$ or $b$ is put in and hence no longer available as a right end for a path.  This gives two possible configurations illustrated in the bottom two parts of Figure~\ref{fig messy}.

  Now we know that $I_{n-2}^{W}  = (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}$ so the paths for building plusses from $I_{n-2}^{W}$ go from the set $\{n-5, n-3\}$ along with whichever of $a$ and $b$ is not in $I_{n}^{V}$ to $\{n-2, n-1, n\}$.  This means that we get plusses as in Figure~\ref{fig messyD} where the left and right cases correspond to the left and right cases in the bottom parts of Figure~\ref{fig messy}

@@

  This proves the first item of statement of the lemma.

  Now consider $I_{n-3}^{V}$.  By \eqref{eq necklace} $I_{n-3}^{V}$ contributes the same plusses as $I_{n_2}^{V}$ except that it contributes an $n-5\rightarrow n-3$ plus in place of the $n-5\rightarrow n-2$ plus.  Also, we have $n-3\in I_{n-3}^{V}$ be the location of $q$ and so $I_{k}^{W} = I_k^{V}\cup\{n-2\}$.  Thus the paths for $I^{W}_{n-3}$ are the same as for $I^{V}_{n-3}$ except that the path that did go to $n-3$ now goes to $n-2$.  This cannot conflict with another path since \eqref{eq necklace} shows that $n-2$ only appears in $I_{n-2}^{V}$ among the necklace elements of $V$.

  Also note that $I_{n-3}^{V}, I_{n-2}^{V}$, and $I_{n-1}^{V}$ share their plusses outside of the $n-3$ and $n-2$ columns.  This proves the remaining statements of the lemma except the furthermore.

  Finally, suppose there were a $n-5\rightarrow n-1$ plus in the Le diagram of $V$.  By the algorithm, it would have to come when $n-5 \not\in I_j^{V}$.  By \eqref{eq necklace} this means that it would have to come from $I_{n-4}^{V}$, $I_{n-3}^{V}$, or $I_{n-2}^{V}$. The analysis above shows it does not come from $I_{n-3}^{V}$ or $I_{n-2}^{V}$.   Now, $n-4\in I_{n-4}^{V}$ by the location of $q$ and so $I_{n-4}^{V}$ must give an $n-5\rightarrow n-4$ plus and so cannot give an $n-5\rightarrow n-1$ plus.

\end{proof}

\note{This is a reorganised version of the previous lemma, keeping both in place for now for comparison.}
\begin{lem}\label{lem n-2 bad v2}
Let $V$ and $W$ be as in Lemma~\ref{lem I}, and suppose that $n-2,n-1 \in I_{n-2}^{V}$. Then:
  \begin{enumerate}
  \item $I_{n-2}^W$ contributes an $(n-3)\rightarrow (n-2)$ plus and an $(n-5)\rightarrow (n-1)$ plus, along with any plusses contributed by $I_n^V$.
  \item $I_{n-3}^W$ contributes the same plusses as $I_{n-3}^V$, except that the plus in the $n-3$ column is shifted one square left into the $n-2$ column.
  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $(n-5)\rightarrow (n-2)$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields the $(n-5)\rightarrow (n-1)$ plus from point (1) instead.
  \item All other plusses from $I_{n-2}^{V}$ and all plusses from $I_{n-1}^V$ were already contributed by $I_{n-3}^V$, so they yield no new information.
  \item No element of the Grassmann necklace of $V$ contributes a $(n-5)\rightarrow (n-1)$ plus.
  \end{enumerate} 
\end{lem}

\begin{proof}
  By Lemma~\ref{lem I} we have $I_{n-2}^{W}= (I_{n}^{V} \setminus \{n-5\})\cup \{n-1,n-2\}$ and $I_1^W = I_1^V\cup \{n-3\}$. We are necessarily in the case where $W$ admits a propagator $q$ (the right hand case of Figure~\ref{fig special p}) and we can see from the location of $q$ that $n-2\not\in I_{n-3}^{V}$ and that $n-5$ is the index of the lowest vertical edge in $\mathcal{B}$.  Thus this section of the Gramann necklace of $V$ looks like

  \begin{equation}\label{eq necklace}
  I_{n-3}^{V} \xrightarrow[\substack{n-3\text{ out}\\n-2\text{ in}}]{} I_{n-2}^{V} \xrightarrow[\substack{n-2\text{ out}\\n-5\text{ in}}]{} I_{n-1}^{V} \xrightarrow[\substack{n-1\text{ out}\\\text{something in}}]{} I_{n}^{V}  \xrightarrow[\substack{n \text{ out}\\\text{something in}}]{\text{\ no change, {\em or}\ }} I_1^{V}
  \end{equation}
  where the first ``something'' is either $n$ or an element of $I_1^{V}$, and the second ``something'' (if it exists) is an element of $I_1^{V}$. If $n\not\in I_n^V$ then $I_n^V = I_1^V$.

  Using this information we can determine the symmetric difference of $I_{n-2}^{V}$ and $I_1^{V}$, which has size $\leq 3$ by the definition of a Grassmann necklace. We have $n-1, n-2 \in I_{n-2}^{V} \setminus I_1^V$ for certain, and the third element (if it exists) must be $n$. On the other hand we have $n-5 \in I_1^{V} \setminus I_{n-2}^V$, along with at least one and at most two other elements.  If there is one such element call it $a$.  If there are two call them $a$ and $b$ with $a>b$.

  \note{Remove the $I_n^V$ diagrams}

  With this information in hand, we can now tackle the points in the statement of the lemma. We know that 
  \[I_{n-2}^{W}  = (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}, \quad n-5 \in I_1^W, \quad n-3\not\in I_{n-2}^W,\]
 and so the paths for building plusses from $I_{n-2}^{W}$ start at
 \[I_1^W \setminus I_{n-2}^W = \{n-3,n-5\} \cup (I_1^V \setminus I_n^V)\]
 and end at
 \[I_{n-2}^W \setminus I_1^W = \{n-2,n-1\}\cup(I_n^V \setminus I_1^V).\]
This means that $I_{n-2}^W$ contributes plusses as in Figure~\ref{fig messyD}, which proves point (1) and most of point (3). From equation \eqref{eq necklace} we see that $n-2$ appears only in $I_{n-2}^V$, so no other Grassmann necklace term for $V$ can contribute a plus in the $n-2$ column; this completes the proof of item (3).

\note{Modify this figure to include the case where there's no plus for column $n$.}

>>>>

These $C(W)$ parametrize a subspace of $\Gr(k, n)$ as shown in \cite{wilsonloop}, call it $\Sigma(W)$. The Wilson loop diagrams also define a volume form on $\Sigma(W)$: \bas \Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V_{p_r}} \textrm{d}c_{p_r}}{R(W)} \;. \eas The denominator $R(W)$ is a polynomial defined by $2 \times 2$ and $1 \times 1$ minors of $C(W)$ as defined below.

@@

%\begin{lem}\label{lem n-2 bad}
%Let $V$ and $W$ be as in Lemma~\ref{lem I}, and suppose that $n-2,n-1 \in I_{n-2}^{V}$.
%Then $I_{n-2}^{W}$ and $I_{n-3}^{W}$ contribute the following plusses to the Le diagram of $W$
%  \begin{itemize}
%  \item An $(n-3)\rightarrow (n-2)$ plus and an $(n-5)\rightarrow (n-1)$ plus.
%  \item All of the $I_{n-1}^{V}$ plusses.
%  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $n-5\rightarrow n-2$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields an $n-5\rightarrow n-1$ plus instead.
%  \item All other plusses of $I_{n-2}^{V}$.
%  \item In the Le diagram of $V$, $I_{n-3}^{V}$ gives a plus in the $n-3$ column.  This $+$ is shifted over into the $n-2$ column in $W$.
%  \item All other plusses of $I_{n-3}^{V}$.
%  \end{itemize}
%Furthermore, no element of the Gramann necklace of $V$ gives an $n-5\rightarrow n-1$ plus.
%  By Lemma~\ref{lem I} $I_{n-2}^{W}= (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}$.  Also, by the location of $q$ in the WLD,  $n-2\not\in I_{n-3}^{V}$ and and $n-5$ is the index of the lowest vertical edge in $\mathcal{B}$.  Thus this section of the Gramann necklace of $V$ looks like
%  \begin{equation}\label{eq necklace}
%  I_{n-3}^{V} \underset{\substack{n-3\text{ out}\\n-2\text{ in}}}{\rightarrow} I_{n-2}^{V} \underset{\substack{n-2\text{ out}\\n-5\text{ in}}}{\rightarrow} I_{n-1}^{V} \underset{\substack{n-1\text{ out}\\\text{something in}}}{\rightarrow} I_{n}^{V} \underset{\substack{n \text{ out}\\\text{something in}}}{\rightarrow} I_1^{V}
%  \end{equation}
%  where the first ``something'' is either $n$ or an element of $I_1^{V}$ and the second ``something'' is an element of $I_1^{V}$.  Additionally all elements not explicitly mentioned must be in $I_1^{V}$ as they remain unchanged through this portion of the necklace.
%  Using this information now determine the symmetric difference of $I_{n-2}^{V}$ and $I_1^{V}$: $n-1, n-2$ and possibly $n$ are in $I_{n-2}^{V}$ but not in $I_1^{V}$.  $n-5$ is in $I_1^{V}$ as are at least one and at most two other elements.  If there is one such element call it $a$.  If there are two call them $a$ and $b$ with $a>b$.  This means that the plusses in the Le diagram of $V$ coming from $I_{n-2}^{V}$ are as in the first part of Figure~\ref{fig messy}.  Stepping to $I_{n-1}^{V}$ simply removes the $n-5\rightarrow n-2$ path, see the second part of Figure~\ref{fig messy}.
%  \begin{figure}
%    \includegraphics{messy}
%    \caption{Plusses coming from $I_{n-2}^{V}$ (top left),  $I_{n-1}^{V}$ (top right) and $I_{n}^{V}$ bottom when $n-1, n-2\in I_{n-1}^{V}$.  The blue lines are the non-intersecting paths.  The dashed blue lines may or may not appear, but if one appears then they both do.}\label{fig messy}
%  \end{figure}
%  Stepping to $I_{n}^{V}$, $n-1$ is taken out and either $n$ is put in if it was not there before, or one of $a$ or $b$ is put in and hence no longer available as a right end for a path.  This gives two possible configurations illustrated in the bottom two parts of Figure~\ref{fig messy}.
%  Now we know that $I_{n-2}^{W}  = (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}$ so the paths for building plusses from $I_{n-2}^{W}$ go from the set $\{n-5, n-3\}$ along with whichever of $a$ and $b$ is not in $I_{n}^{V}$ to $\{n-2, n-1, n\}$.  This means that we get plusses as in Figure~\ref{fig messyD} where the left and right cases correspond to the left and right cases in the bottom parts of Figure~\ref{fig messy}
%  \begin{figure}
%     \includegraphics{messyD}
%    \caption{Plusses coming from $I_{n-2}^{W}$.}\label{fig messyD}
%  \end{figure}
%  This proves the first item of statement of the lemma.
%  Now consider $I_{n-3}^{V}$.  By \eqref{eq necklace} $I_{n-3}^{V}$ contributes the same plusses as $I_{n_2}^{V}$ except that it contributes an $n-5\rightarrow n-3$ plus in place of the $n-5\rightarrow n-2$ plus.  Also, we have $n-3\in I_{n-3}^{V}$ be the location of $q$ and so $I_{k}^{W} = I_k^{V}\cup\{n-2\}$.  Thus the paths for $I^{W}_{n-3}$ are the same as for $I^{V}_{n-3}$ except that the path that did go to $n-3$ now goes to $n-2$.  This cannot conflict with another path since \eqref{eq necklace} shows that $n-2$ only appears in $I_{n-2}^{V}$ among the necklace elements of $V$.
%  Also note that $I_{n-3}^{V}, I_{n-2}^{V}$, and $I_{n-1}^{V}$ share their plusses outside of the $n-3$ and $n-2$ columns.  This proves the remaining statements of the lemma except the furthermore.
%  Finally, suppose there were a $n-5\rightarrow n-1$ plus in the Le diagram of $V$.  By the algorithm, it would have to come when $n-5 \not\in I_j^{V}$.  By \eqref{eq necklace} this means that it would have to come from $I_{n-4}^{V}$, $I_{n-3}^{V}$, or $I_{n-2}^{V}$. The analysis above shows it does not come from $I_{n-3}^{V}$ or $I_{n-2}^{V}$.   Now, $n-4\in I_{n-4}^{V}$ by the location of $q$ and so $I_{n-4}^{V}$ must give an $n-5\rightarrow n-4$ plus and so cannot give an $n-5\rightarrow n-1$ plus.

\begin{lem}\label{lem n-2 bad}
Let $V$ and $W$ be as in Lemma~\ref{lem I}, and suppose that $n-2,n-1 \in I_{n-2}^{V}$. Then:
  \begin{enumerate}
  \item $I_{n-2}^W$ contributes an $(n-3)\rightarrow (n-2)$ plus and an $(n-5)\rightarrow (n-1)$ plus, along with any plusses contributed by $I_n^V$.
  \item $I_{n-3}^W$ contributes the same plusses as $I_{n-3}^V$, except that the plus in the $n-3$ column (which necessarily exists) is shifted one square left into the $n-2$ column.
  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $(n-5)\rightarrow (n-2)$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields the $(n-5)\rightarrow (n-1)$ plus from point (1) instead.
  \item All other plusses from $I_{n-2}^{V}$ and all plusses from $I_{n-1}^V$ were already contributed by $I_{n-3}^V$, so they yield no new information.
  \item No element of the Grassmann necklace of $V$ contributes a $(n-5)\rightarrow (n-1)$ plus.
  \end{enumerate} 
\end{lem}

\begin{proof}
  By Lemma~\ref{lem I} we have $I_{n-2}^{W}= (I_{n}^{V} \setminus \{n-5\})\cup \{n-1,n-2\}$ and $I_1^W = I_1^V\cup \{n-3\}$. We are necessarily in the case where $W$ admits a propagator $q$ (the right hand case of Figure~\ref{fig special p}) and we can see from the location of $q$ that $n-2\not\in I_{n-3}^{V}$ and that $n-5$ is the index of the lowest vertical edge in $\mathcal{B}$.  Also note that by Lemma~\ref{vertex cyclic int lem} and Corollary~\ref{no coloops} we must have $n\in I_n^V$. Thus this section of the Gramann necklace of $V$ looks like

  \begin{equation}\label{eq necklace}
  I_{n-3}^{V} \xrightarrow[\substack{n-3\text{ out}\\n-2\text{ in}}]{} I_{n-2}^{V} \xrightarrow[\substack{n-2\text{ out}\\n-5\text{ in}}]{} I_{n-1}^{V} \xrightarrow[\substack{n-1\text{ out}\\\text{something in}}]{} I_{n}^{V}  \xrightarrow[\substack{n \text{ out}\\\text{something in}}]{} I_1^{V}
  \end{equation}
  where the first ``something'' is either $n$ or an element of $I_1^{V}$, and the second ``something'' is an element of $I_1^{V}$.

  Using this information we can determine the symmetric difference of $I_{n-2}^{V}$ and $I_1^{V}$.  Note that $I_{n-2}^V\setminus I_1^V$ and $I_1^V\setminus I_{n-2}^V$ each have size $\leq 3$ by the definition of a Grassmann necklace. We have $n-1, n-2 \in I_{n-2}^{V} \setminus I_1^V$ for certain, and the third element (if it exists) must be $n$. On the other hand we have $n-5 \in I_1^{V} \setminus I_{n-2}^V$, along with at least one and at most two other elements.  If there is one such element call it $a$.  If there are two call them $a$ and $b$ with $a>b$.

  With this information in hand, we can now tackle the points in the statement of the lemma. We know that 
  \[I_{n-2}^{W}  = (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}, \quad n-5 \in I_1^W, \quad n-3\not\in I_{n-2}^W,\]
 and so the paths for building plusses from $I_{n-2}^{W}$ start at
 \[I_1^W \setminus I_{n-2}^W = \{n-3,n-5\} \cup (I_1^V \setminus I_n^V)\]
 and end at
 \[I_{n-2}^W \setminus I_1^W = \{n-2,n-1\}\cup(I_n^V \setminus I_1^V).\]
This means that $I_{n-2}^W$ contributes plusses as in Figure~\ref{fig messyD}, while $I_n^V$ contributes as in the bottom line of Figure~\ref{fig messy}, which proves point (1) and most of point (3). From equation \eqref{eq necklace} we see that $n-2$ appears only in $I_{n-2}^V$, so no other Grassmann necklace term for $V$ can contribute a plus in the $n-2$ column; this completes the proof of item (3).

@@
